---
title: "WUP_04"
author: "Syslee Evans"
date: "Summer 2022"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)

```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

```{r}
gssDataframe=read.csv("GSScsv.csv")
```
WORDSUM is a numerical variable,DEGREE is a catagorical variable. 
# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results


<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
boxplot(WORDSUM~DEGREE,data=gssDataframe)
```

From the boxplot it appears that the WORDSUMS are different because the students in less than highschool have a medium less than that the studernts in high school and the students in bachlors.


### Numerical Descriptive Results
```{r}
favstats(WORDSUM~DEGREE,data=gssDataframe)
```
For this table it seems like the median of the graduate students is the greatest with an 11. it looks like theres a difference in the ability in the WORDSUM catagory. 
<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

## Inferential Results
```{r}
model1 <- aov(WORDSUM~DEGREE,data=gssDataframe)
summary.aov(model1)
```
The Pvalue is 0.000236.
Because the pvalue is less than 0.05 we reject the null hypothisis. 
```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 8.7, 2, 2))
plot(TukeyHSD(model1),las=2)
```
based on the graph, there is a difference between high school graduates and also betrween less than and graduate.
this is because neither one of them cross or touch the zero level. 